Solar wind predictions for the Parker Solar Probe orbit

### 2017-08-23 paperMVVB/sections/content.tex
fig 7+10, summary bullets

%solar wind parameters
It is not known which specific solar wind type or structure PSP will encounter at a given time during its mission, therefore we extrapolate the probability distributions of the major solar wind parameters from existing solar wind measurements and take solar cycle dependencies into account.
As a baseline we describe the solar wind environment through the key quantities of a magnetized plasma: \textit{density}, \textit{temperature} and \textit{magnetic field strength}. Furthermore, the bulk flow \textit{velocity} is the defining parameter of the two types of solar wind. Solar wind quantities, like flux densities, mass flux and plasma beta, can directly be derived from these four parameters.

@@

The median and mean values of the key solar wind parameters for different solar distances of the Helios data are calculated for the minimal distance resolution \SI{0.01}{\au} of the data set, see Fig.~\ref{fig:radial_fit_4_thesis_light_skip_pdfcairo_plot}. Assuming a radial solar wind outflow, it is expected that the distance dependence of the solar wind parameters over the Helios data range \SIrange{0.29}{0.98}{\au} can be described through power law scaling. Therefore we use the power law function
\begin{align}
	x(r) = d\cdot r^e	\label{eq:power_function}
\end{align}
for the regression fit of the median and mean, with $r$ being the solar distance in astronomical units, $d$ the magnitude at \SI{1}{\au} and $e$ the exponent. The fits are weighted through the different data counts per bin.
%fit result table and figure
The fit coefficients ($d_\text{med}$, $d_\text{avg}$, $e_\text{med}$ and $e_\text{avg}$) are listed in Table~\ref{tab:mean_median_fit_parameter}.
\begin{table*}
	\caption{Fit coefficients for the median and mean solar distance dependencies of the four solar wind parameters derived from the combined Helios~1 and 2 data. The errors in brackets are the estimated standard deviations of each fit parameter. The crossing distances indicate where the median and mean fits intersect each other. The yearly variation is the weighted standard deviation derived from the yearly fit exponents seen in Fig.~\ref{fig:yearly_gradients_c}.}
	\label{tab:mean_median_fit_parameter}
	\centering
	\begin{tabular}{l
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	@{}c
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	S[table-format = 1.1(2)e1]
	S[table-format = 1.3]}
		\hline\hline
		\multirow{2}{*}{Parameter}	&\multicolumn{2}{c}{Median}	&	&\multicolumn{2}{c}{Mean}	&\multicolumn{1}{c}{Crossing distance}	&\multicolumn{1}{c}{Yearly variation}\\
		\cline{2-3}	\cline{5-6}
			&\multicolumn{1}{c}{$d_\text{med}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{med}$}	&	&\multicolumn{1}{c}{$d_\text{avg}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{avg}$}	&\multicolumn{1}{c}{[\si{\au}]}	&\multicolumn{1}{c}{$\Delta e$}\\
		\hline
		Magnetic field	&5.377(92)	&-1.655(17)	&	&6.05(10)	&-1.546(18)	&0.339(11)	&0.11\\
		Velocity	&4.107(28)	&0.058(13)	&	&4.356(24)	&0.049(10)	&0.7(83)e3	&0.012\\
		Density		&5.61(27)	&-2.093(46)	&	&7.57(30)	&-2.010(38)	&0.027(73)	&0.072\\
		Temperature	&7.14(23)	&-0.913(39)	&	&9.67(21)	&-0.792(28)	&0.082(85)	&0.050\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{In units of \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}
% source year variation: fit_out_yearly_gradients_table_b_stats.txt

@@

%crossing distances
Fig.~\ref{fig:radial_fit_4_thesis_light_skip_pdfcairo_plot} shows the radial dependence of the solar wind parameters over the distance range \SIrange{0.29}{0.98}{\au} and the mean and median values and their respective power law fits. The mean and median velocity fit exponents are very similar, which indicates that they just as well can be kept identical so that the basic shape of the frequency distribution does not change with distance.
\begin{figure*}
	\includegraphics[width=18cm]{figures/radial_fit_4_thesis_light_skip_pdfcairo_plot.pdf}
	\caption{Helios hourly data plots of the four solar wind parameters over solar distance. The mean and median per \SI{0.01}{\au} data bin and their fit curves are plotted as well. The Helios data has a native distance resolution of \SI{0.01}{\au}, thus, to make the abundance visible in these plots, we added a random distance value of up to \SI{+-0.005}{\au}.}
	\label{fig:radial_fit_4_thesis_light_skip_pdfcairo_plot}
\end{figure*}
Contrary, the mean and median fits for the magnetic field strength cross each other at \SI{0.339}{\au} and the mean is lower than the median at smaller distances (Table~\ref{tab:mean_median_fit_parameter}). Thus, below that distance the distribution function cannot well be described anymore by a lognormal function. The fits for the proton temperature show a similar behavior,
having an intersection at \SI{0.082}{\au}. Therefore the extrapolation of the magnetic field and temperature distribution frequencies to the PSP orbit by applying lognormal functions is limited. To circumvent such limitations we set the exponents $e_\text{med}$ and $e_\text{avg}$ to be identical for all four parameters. It should be noted that this simplification leads to slightly larger modeling errors, especially in case of the magnetic field strength.

@@

We extrapolate the derived empirical solar wind model (\ref{eq:general_sw_model}) to PSP’s orbital distance range and compare the results with those from the existing models shown in Fig.~\ref{fig:sw_extrapolation_ssn_b_plot}.
\begin{figure*}
	\includegraphics[width=18cm]{figures/sw_extrapolation_ssn_b_plot.pdf}
	\caption{Radial extrapolation of the solar wind parameters to the PSP orbit region. The from Helios and OMNI measurements obtained models are extrapolated to the PSP region---for the extreme cases of solar minimum (SSN = 0) and maximum (SSN = 200). Note that there is a time lag to the SSN depending on the solar wind parameter. The magnetic field radial dependence is slightly flatter than the analytic DQCS model for solar minimum which \citet{Banaszkiewicz1998} derived. \hl{Below} \SI{20}{\Rs} the slow wind velocity is overestimated in comparison to the measurements from \citet{Wang2000}) and \citet{Sheeley1997}. Down to PSP's perihelion the density is in good agreement with the model from \citet{Leblanc1998}. to 1-column...?}
	\label{fig:sw_extrapolation_ssn_b_plot}
\end{figure*}

%magnetic field:
The magnetic field strength is found to increase from median values of about \SI{43}{\nT} at \SI{0.25}{\au} to \SI{715}{\nT} at \SI{0.046}{\au} for a SSN of 0. Taking a SSN of 200 increases the value to \SI{69}{\nT} at \SI{0.25}{\au} and \SI{1152}{\nT} at \SI{0.046}{\au}. Our extrapolation results are slightly flatter than those derived from the analytical magnetic field model by \citet{Banaszkiewicz1998}, who constructed a dipole plus quadrupole plus current sheet (DQCS) model.
This difference is arguably due to the previously mentioned (Sect.~\ref{sec:solar_distance_dependency}) with solar distance changing shape of the frequency distribution, which for smaller distances deviates more from the model’s lognormal distribution.

%velocity:
The average velocity is found to decrease from \SI{340}{\km\per\s} at \SI{0.25}{\au} to about \SI{290}{\km\per\s} at \SI{0.046}{\au} for a SSN of 0. Whereas using a SSN of 200 it decreases from \SI{390}{\km\per\s} to \SI{330}{\km\per\s}. Comparing the results with those found by \citet{Sheeley1997} and \citet{Wang2000} shows an overestimation in our extrapolated velocity values for distances \hl{below} \SI{20}{\Rs}. They used LASCO coronagraph observations to track moving coronal features (blobs) in the distance range \SIrange{2}{30}{\Rs} to determine speed profiles and sources of the slow solar wind and they derived temperature and sonic point values for slow solar wind with the isothermal expansion model \citep{Parker1958}. Therefore, it generally can be expected that PSP will encounter a slower solar wind environment close to the Sun than our model estimates and thus PSP will measure solar wind acceleration processes \hl{below} \SI{30}{\Rs} \citep{Sheeley1997,McComas2008}.

@@

%temperature:
The extrapolated proton temperature increases from about \SI{260000}{\K} at \SI{0.25}{\au} to about \SI{1690000}{\K} at \SI{0.046}{\au} for a SSN of 0 and from \SI{440000}{\K} to \SI{2860000}{\K} for a SSN of 200. Knowing that near-Sun coronal temperatures are in the range of \SIrange{2}{3}{\mega\K} \citep{Billings1959,Liebenberg1975}, the model \hl{may overestimate} the extrapolated temperatures at the PSP perihelion distance. {\color{red} The results can be compared to...}

@@

We started the development of the empirical solar wind environment model for the near-ecliptic PSP orbit by lognormally fitting the about 40~years of in situ near-Earth solar wind data collected in the OMNI database, using the frequency distributions of the key solar wind parameters magnetic field strength, velocity, density and temperature. The OMNI multi-satellite solar wind data is intercalibrated, covers almost five solar cycles and thus represents solar wind gathered at different phases of solar activity in the ecliptic plane. In the next step we investigated the yearly variation of the solar wind distribution functions along with the SSN over 53~years and derived linear dependencies of the solar wind parameters with the SSN. The radial dependencies of the solar wind distribution functions were then analyzed using Helios~1 and 2 data for the distance range \SIrange{0.29}{0.98}{\au} in bins of \SI{0.01}{\au}. The herefrom derived power law fit functions were used to scale the prior calculated SSN-dependent \SI{1}{\au} distribution fit functions to the PSP orbit, combined with SSN predictions for the years 2018--2025, i.e., for the prime mission phase that gets PSP to closest perihelia of 9.86~solar radii distance to the Sun. The reason for performing the analysis this way is based on the fact that the OMNI solar wind database is much larger than the Helios database. Though, it is clear that the calculated distribution functions just represent first order estimates of the real solar wind to be encountered by PSP. The solar wind environment to be encountered will depend at times of PSP on the structure of the solar corona and underlying photospheric magnetic field {\color{red} , as shown by \hl{xxx (ref.)},} and on the evolution and interaction of individual solar wind streams and superimposed CMEs and shocks.

However, the derived results are in good agreement with existing studies about near-Sun solar wind magnetic field strenghts and densities as shown in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}. The from direct measurements differing velocity and temperature extrapolations below \hl{20~solar radii} indicate that PSP will indeed dive into the acceleration and heating regions of the solar wind to be expected at these distances (see Fig.~\ref{fig:sw_extrapolation_ssn_b_plot}). The near-Sun solar wind velocity at PSP perihelion is also expected to be slower than our model estimates, because the region of the Alfvénic critical surface, up to which the solar wind is believed to be accelerated, is predicted to lie in average around \SI{17}{\Rs} \citep[e.g.,][]{Sittler1999,Exarhos2000} and correlates with solar activity between \SI{15}{\Rs} in minimum and \SI{30}{\Rs} in maximum \citep{Katsikas2010,Goelzer2014}.	% in average around 17~Rs (e.g., 0.6~Rs Schatten 1969, 17~Rs Sittler1999, 17~Rs Exarhos2000) and scales with solar activity between 15~Rs in minimum and 30~Rs in maximum (19~Rs Katsikas2010, 15–30~Rs Goelzer2014).

In our study we have not specifically investigated the occurrences of extreme solar wind parameters caused by CMEs or enhanced values due to stream interaction or co-rotating interaction regions. The Helios solar wind measurements plotted over radial distance in \hl{Fig.~}\ref{fig:radial_fit_4_thesis_light_skip_pdfcairo_plot} show several extreme values far above the usual solar wind velocities likely to be associated with individual CMEs. The results by \citet{Sachdeva2017} indicate that due to the prevailing of solar wind drag the speeds of fast CMEs will commonly slow down substantially from early distances of a few solar radii. Therefore, it is expected that PSP will encounter CMEs with much higher speeds than those observed during the Helios mission. Also, the magnetic field, density and temperature values are expected to be much larger than in the average solar wind in individual fast shock associated CME events. PSP will thus also substantially improve our understanding of the near-Sun evolution of CMEs and their expansion with radial distance.

a\\

note special velocity treatment with double lognormal fits\\
slow/fast streams basically maintain characteristic speeds with solar cycle\\

@@

slow/fast ratio SSN dependency\\
individual velocity part discussion -> there is no specific velocity threshold between slow and fast solar wind types, the velocity ranges of both types overlap.\\
Not only the slowest wind but also the fastest wind is expected to converge to the average speed (Sanchez-Diaz2016 p.~2835, using MHD-model -> very slow solar wind is continuation of slow wind) (because of interaction).\\
The ratio of both varies with solar activity, e.g., 3~years after maximum, polar coronal holes are observed to often have equatorial extensions (cite?). see and use \citet{Bougeret1984} p.~498...\\
for the overlapping velocity model the SSN dependency is steeper than for a simple threshold\\

\section{Summary of the results}
For the near-Earth solar wind OMNI data and the Helios~1 and 2 data, obtained over the distance range \SIrange{0.29}{0.98}{\au}, we derived lognormal representations of the frequency distributions’ shapes of the four key solar wind parameters magnetic field strength, proton velocity, density and proton temperature. The dependencies of these frequency distributions on the solar activity cycle and on radial distance to the Sun have been modeled with analytical relations and extrapolated to the Parker Solar Probe orbit, taking into account predictions of the sunspot number. This empirical CGAUSS solar wind model for PSP, representing the solar wind’s solar activity and distance behavior, yields the following main results for the bulk solar wind:\\

{\color{red}
parameter relationships as bullets...\\

Bmed(SSN,r) = ...\\
V...\\
give valid ranges [a--b]...\\
We estimated solar wind median values during PSP’s first perihelion and the values for PSP’s first minimum perihelion of 9.86~Rs in 2024.\\
}

The velocity and temperature values are overestimated below \SI{20}{\Rs} in comparison with existing observations. This shows that solar wind acceleration and heating processes below \SI{20}{\Rs} limit the simple back-extrapolation from existing in situ measurements.\\

%Implications for WISPR mission operations and SoloHI on Solar Orbiter...
The knowledge of the expected solar wind environment helps to optimize the WISPR and PSP preplanning of the science operations. This also applies for the Heliospheric Imager (SoloHI) onboard the Solar Orbiter spacecraft (Howard2013).\\

>>>>

%solar wind parameters
It is not known which specific solar wind type or structure PSP will encounter at a given time during its mission, therefore we extrapolate the probability distributions of the major solar wind parameters from existing solar wind measurements and take solar cycle dependencies into account.
As a baseline we describe the solar wind environment through the key quantities of a magnetized plasma: \textit{magnetic field strength}, \textit{density} and \textit{temperature}. Furthermore, the bulk flow \textit{velocity} is the defining parameter of the two types of solar wind. Solar wind quantities, like flux densities, mass flux and plasma beta, can directly be derived from these four parameters. In the analyses we treat the solar wind as a proton plasma---the average helium abundance is about \SI{4.5}{\percent} and in slow wind at solar cycle minimum even less than \SI{2}{\percent} \citep{Feldman1978,Schwenn1983,Kasper2012}.

@@

The median and mean values of the key solar wind parameters for different solar distances of the Helios data are calculated for the minimal distance resolution \SI{0.01}{\au} of the data set, see Fig.~\ref{fig:radial_fit_4_thesis_light_b_skip}. Assuming a radial solar wind outflow, it is expected that the distance dependence of the solar wind parameters over the Helios data range \SIrange{0.29}{0.98}{\au} can be described through power law scaling. Therefore we use the power law function
\begin{align}
	x(r) = d\cdot r^e	\label{eq:power_function}
\end{align}
for the regression fit of the median and mean, with $r$ being the solar distance in astronomical units, $d$ the magnitude at \SI{1}{\au} and $e$ the exponent. The fits are weighted through the different data counts per bin.
%fit result table and figure
The fit coefficients ($d_\text{med}$, $d_\text{avg}$, $e_\text{med}$ and $e_\text{avg}$) are listed in Table~\ref{tab:mean_median_fit_parameter}.
\begin{table*}
	\caption{Fit coefficients for the median and mean solar distance dependencies of the four solar wind parameters derived from the combined Helios~1 and 2 data. The errors in brackets are the estimated standard deviations of each fit parameter. The crossing distances indicate where the median and mean fits intersect each other. The yearly variation is the weighted standard deviation derived from the yearly fit exponents seen in Fig.~\ref{fig:yearly_gradients_c}.}
	\label{tab:mean_median_fit_parameter}
	\centering
	\begin{tabular}{l
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	@{}c
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	S[table-format = 1.1(2)e1]
	S[table-format = 1.3]}
		\hline\hline
		\multirow{2}{*}{Parameter}	&\multicolumn{2}{c}{Median}	&	&\multicolumn{2}{c}{Mean}	&\multicolumn{1}{c}{Crossing distance}	&\multicolumn{1}{c}{Yearly variation}\\
		\cline{2-3}	\cline{5-6}
			&\multicolumn{1}{c}{$d_\text{med}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{med}$}	&	&\multicolumn{1}{c}{$d_\text{avg}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{avg}$}	&\multicolumn{1}{c}{[\si{\au}]}	&\multicolumn{1}{c}{$\Delta e$}\\
		\hline
		Magnetic field	&5.377(92)	&-1.655(17)	&	&6.05(10)	&-1.546(18)	&0.339(11)	&0.11\\
		Velocity	&4.107(28)	&0.058(13)	&	&4.356(24)	&0.049(10)	&0.7(83)e3	&0.012\\
		Density		&5.61(27)	&-2.093(46)	&	&7.57(30)	&-2.010(38)	&0.027(73)	&0.072\\
		Temperature	&7.14(23)	&-0.913(39)	&	&9.67(21)	&-0.792(28)	&0.082(85)	&0.050\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{In units of \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}
% source year variation: fit_out_yearly_gradients_table_b_stats.txt

@@

%crossing distances
Fig.~\ref{fig:radial_fit_4_thesis_light_b_skip} shows the radial dependence of the solar wind parameters over the distance range \SIrange{0.29}{0.98}{\au} and the mean and median values and their respective power law fits. The mean and median velocity fit exponents are very similar, which indicates that they just as well can be kept identical so that the basic shape of the frequency distribution does not change with distance.
\begin{figure*}
	\includegraphics[width=18cm]{figures/radial_fit_4_thesis_light_b_skip.pdf}
	\caption{Helios hourly data plots of the four solar wind parameters over solar distance. The mean and median per \SI{0.01}{\au} data bin and their fit curves are plotted as well. The Helios data has a native distance resolution of \SI{0.01}{\au}, thus, to make the abundance visible in these plots, we added a random distance value of up to \SI{+-0.005}{\au}. The maximal velocities of CME events larger than \SI{800}{\km\per\s} are indicated with circles.}
	\label{fig:radial_fit_4_thesis_light_b_skip}
\end{figure*}
Contrary, the mean and median fits for the magnetic field strength cross each other at \SI{0.339}{\au} and the mean is lower than the median at smaller distances (Table~\ref{tab:mean_median_fit_parameter}). Thus, below that distance the distribution function cannot well be described anymore by a lognormal function. The fits for the proton temperature show a similar behavior,
having an intersection at \SI{0.082}{\au}. Therefore the extrapolation of the magnetic field and temperature distribution frequencies to the PSP orbit by applying lognormal functions is limited. To circumvent such limitations we set the exponents $e_\text{med}$ and $e_\text{avg}$ to be identical for all four parameters. It should be noted that this simplification leads to slightly larger modeling errors, especially in case of the magnetic field strength.

@@

We extrapolate the derived empirical solar wind model (\ref{eq:general_sw_model}) to PSP’s orbital distance range and compare the results with those from the existing models shown in Fig.~\ref{fig:sw_extrapolation_ssn_d_plot}.
\begin{figure*}
	\includegraphics[width=18cm]{figures/sw_extrapolation_ssn_d_plot.pdf}
	\caption{Radial extrapolation of the solar wind median parameters to the PSP orbit region. The from Helios and OMNI measurements obtained models are extrapolated to the PSP region---for the extreme cases of solar minimum (SSN = 0) and maximum (SSN = 200). Note that there is a time lag to the SSN depending on the solar wind parameter. Also plotted are the radial dependencies of the analytic DQCS magnetic field model for solar minimum from \citet{Banaszkiewicz1998}, the slow wind velocity models from \citet{Wang2000} and \citet{Sheeley1997}, the density model from \citet{Leblanc1998} and the range of temperature measurements from \citet{Billings1959} and \citet{Liebenberg1975}.}
	\label{fig:sw_extrapolation_ssn_d_plot}
\end{figure*}

%magnetic field:
The magnetic field strength is found to increase from median values of about \SI{43}{\nT} at \SI{0.25}{\au} to \SI{715}{\nT} at \SI{0.046}{\au} for a SSN of 0. Taking a SSN of 200 increases the values to \SI{69}{\nT} and \SI{1152}{\nT}. Our extrapolation results are slightly flatter than those derived from the analytical magnetic field model by \citet{Banaszkiewicz1998}, who constructed an analytic dipole plus quadrupole plus current sheet (DQCS) model for solar minimum.
This difference is arguably due to the previously mentioned (Sect.~\ref{sec:solar_distance_dependency}) with solar distance changing shape of the frequency distribution, which for smaller distances deviates more from the model’s lognormal distribution.

%velocity:
The average velocity is found to decrease from \SI{340}{\km\per\s} at \SI{0.25}{\au} to about \SI{290}{\km\per\s} at \SI{0.046}{\au} for a SSN of 0. Whereas using a SSN of 200 it decreases from \SI{390}{\km\per\s} to \SI{330}{\km\per\s}. Comparing the results with the measurements by \citet{Sheeley1997} and \citet{Wang2000} shows an overestimation in our extrapolated slow solar wind velocity values for distances below about \SI{20}{\Rs}. They used LASCO coronagraph observations to track moving coronal features (blobs) in the distance range \SIrange{2}{30}{\Rs} to determine speed profiles and sources of the slow solar wind and they derived temperature and sonic point values for slow solar wind with the isothermal expansion model from \citet{Parker1958}. Therefore, it generally can be expected that PSP will encounter a slower solar wind environment close to the Sun than our model estimates and thus PSP will measure solar wind acceleration processes \hl{below} \SI{30}{\Rs} \citep{Sheeley1997,McComas2008}.

@@

%temperature:
The extrapolated proton temperature increases from about \SI{260000}{\K} at \SI{0.25}{\au} to about \SI{1690000}{\K} at \SI{0.046}{\au} for a SSN of 0 and from \SI{440000}{\K} to \SI{2860000}{\K} for a SSN of 200. Knowing that near-Sun coronal temperatures are in the range of \SIrange{2}{3}{\mega\K} \citep{Billings1959,Liebenberg1975}, the model overestimate the extrapolated temperatures at the PSP perihelion distance.

@@

We started the development of the empirical solar wind environment model for the near-ecliptic PSP orbit by lognormally fitting the about 40~years of in situ near-Earth solar wind data collected in the OMNI database, using the frequency distributions of the key solar wind parameters magnetic field strength, velocity, density and temperature. The OMNI multi-spacecraft solar wind data is intercalibrated and covers almost five solar cycles. It thus represents solar wind gathered at different phases of solar activity in the ecliptic plane. In the next step we investigated the yearly variation of the solar wind distribution functions along with the SSN over 53~years and derived linear dependencies of the solar wind parameters with the SSN. The radial dependencies of the solar wind distribution functions were then analyzed using Helios~1 and 2 data for the distance range \SIrange{0.29}{0.98}{\au} in bins of \SI{0.01}{\au}. The herefrom derived power law fit functions were used to scale the prior calculated SSN-dependent \SI{1}{\au} distribution fit functions to the PSP orbit, combined with SSN predictions for the years 2018--2025, i.e., for the prime mission phase that gets PSP to closest perihelia of 9.86~solar radii distance. The reason for performing the analysis this way is based on the fact that the OMNI solar wind database is much larger than the Helios database. Though, it is clear that the calculated distribution functions just represent first order estimates of the real solar wind to be encountered by PSP. The solar wind environment to be encountered will depend at times of PSP on the structure of the solar corona and underlying photospheric magnetic field and on the evolution and interaction of individual solar wind streams and superimposed CMEs and shocks.

However, the derived results are in good agreement with existing studies about near-Sun solar wind magnetic field strenghts and densities as shown in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}. The from direct measurements differing velocity and temperature extrapolations below about 20~solar radii indicate that PSP will indeed dive into the acceleration and heating regions of the solar wind to be expected at these distances (see Fig.~\ref{fig:sw_extrapolation_ssn_d_plot}). The near-Sun solar wind velocity at PSP perihelion is also expected to be slower than our model estimates, because the region of the Alfvénic critical surface, up to which the solar wind is believed to be accelerated, is predicted to lie in average around \SI{17}{\Rs} \citep[e.g.,][]{Sittler1999,Exarhos2000} and correlates with solar activity between \SI{15}{\Rs} in cycle minimum and \SI{30}{\Rs} in maximum \citep{Katsikas2010,Goelzer2014}.	% in average around 17~Rs (e.g., 0.6~Rs Schatten 1969, 17~Rs Sittler1999, 17~Rs Exarhos2000) and scales with solar activity between 15~Rs in minimum and 30~Rs in maximum (19~Rs Katsikas2010, 15–30~Rs Goelzer2014).

In our study we have not specifically investigated the occurrences of extreme solar wind parameters caused by CMEs or enhanced values due to stream interaction or co-rotating interaction regions. The Helios solar wind measurements plotted over radial distance in Fig.~\ref{fig:radial_fit_4_thesis_light_b_skip} show several extreme values far above the usual solar wind velocities, likely to be associated with individual CMEs. The results by \citet{Sachdeva2017} indicate that due to the prevailing of solar wind drag the speeds of fast CMEs will commonly slow down substantially from early distances of a few solar radii. Therefore, it is expected that PSP will encounter CMEs with much higher speeds than those observed during the Helios mission. Also, the magnetic field, density and temperature values are expected to be much larger than in the average solar wind in individual fast shock associated CME events. PSP will thus also substantially improve our understanding of the near-Sun evolution of CMEs and their expansion with radial distance.\\

{\color{red}
note special velocity treatment with double lognormal fits\\
slow/fast streams basically maintain characteristic speeds with solar cycle\\

@@

slow/fast ratio SSN dependency\\
individual velocity part discussion -> there is no specific velocity threshold between slow and fast solar wind types, the velocity ranges of both types overlap.\\
Not only the slowest wind but also the fastest wind is expected to converge to the average speed (Sanchez-Diaz2016 p.~2835, using MHD-model -> very slow solar wind is continuation of slow wind) (because of interaction).\\
The ratio of both varies with solar activity, e.g., 3~years after maximum, polar coronal holes are observed to often have equatorial extensions (cite?). see and use \citet{Bougeret1984} p.~498...\\
for the overlapping velocity model the SSN dependency is steeper than for a simple threshold\\
}

\section{Summary}
For the near-Earth solar wind OMNI data and the Helios~1 and 2 data, obtained over the distance range \SIrange{0.29}{0.98}{\au}, we derived lognormal representations of the frequency distributions’ shapes of the four key solar wind parameters magnetic field strength, proton velocity, proton density and proton temperature. The dependencies of these frequency distributions on the solar activity cycle and on radial distance to the Sun have been modeled with analytical relations and extrapolated to the Parker Solar Probe orbit, taking into account predictions of the sunspot number. This empirical CGAUSS solar wind model for PSP, representing the solar wind’s solar activity and distance behavior, yields estimations for the solar wind median values at PSP’s first perihelion in 2018 and at PSP’s closest perihelion in 2024. The model yields the following main results for the bulk solar wind parameters:
\begin{itemize}
	\item The empirical dependency of the magnetic field strength median value on solar activity and radial distance is
	\begin{align}
		B_\text{med}(ssn,r) = (\SI{0.0131}{\nT} \cdot ssn + \SI{4.29}{\nT}) \cdot r^{-1.66}	\,.
	\end{align}
	This approximation appears to be valid for PSP's full distance range, however near its closest perihelion the actual median values might lie within \SI{20}{\%} higher. The estimated median values for PSP's first and closest perihelion are \SI{87}{\nT} and \SI{943}{\nT}.
	\item Averaging over fast and slow solar wind streams, the derived median velocity value dependence is
	\begin{align}
		v_\text{med}(ssn,r) = \left(\SI{1.31}{\km\per\s} \cdot ssn + \SI{429}{\km\per\s}\right) \cdot r^{0.099}	\,.
	\end{align}
	This relation is valid above about \SI{20}{\Rs}, below this solar distance it overestimates the actual velocities obtained from remote measurements. The estimated median values for PSP's first and closest perihelion are \SI{340}{\km\per\s} and \SI{290}{\km\per\s}.
	\item The density relation is found to be
	\begin{align}
		n_\text{med}(ssn,r) = \left(\SI{0.0038}{\per\cm\cubed} \cdot ssn + \SI{4.50}{\per\cm\cubed}\right) \cdot r^{-2.11}	\,.
	\end{align}
	It seems valid throughout the full PSP orbital distance range, even down to about \SI{8}{\Rs}. The estimated median values for PSP's first and closest perihelion are \SI{4015}{\per\cm\cubed} and \SI{9733}{\per\cm\cubed}.
	\item The derived correlation function for the temperature is
	\begin{align}
		T_\text{med}(ssn,r) = (\SI{197}{\K} \cdot ssn + \SI{57300}{\K}) \cdot r^{-1.10}	\,.
	\end{align}
	Near PSP's perihelion the validity of this relation is difficult to evaluate. The estimated median values for PSP's first and closest perihelion are \SI{503000}{\K} and \SI{1930000}{\K}.
\end{itemize}

{\color{red}
The velocity and temperature values are overestimated below \SI{20}{\Rs} in comparison with existing observations. This shows that solar wind acceleration and heating processes below \SI{20}{\Rs} limit the simple back-extrapolation from existing in situ measurements.\\
}

%Implications for WISPR mission operations and SoloHI on Solar Orbiter...
The knowledge of the expected solar wind environment helps to optimize the WISPR and PSP preplanning of the science operations. This also applies for the Heliospheric Imager (SoloHI) onboard the Solar Orbiter spacecraft \citep{Howard2013}.
